---
title: "hw02"
author: "Sierra Park"
date: "9/25/2017"
output: github_document
---


#1 Create a Data Dictionary 
refer to the dictionary file

#2 Import the Data in R

```{r}
setwd('/Users/spark/Documents/UCBERKELEY/2017Fall/stat133/stat133-hws-fall17/hw02/data')
classes <- c("character", "character", "factor", "character", "double", rep("integer",19))
dat <- read.csv(file = "nba2017-player-statistics.csv", colClasses = classes)
dat
```

```{r}
str(dat)
```


```{r}
setwd('/Users/spark/Documents/UCBERKELEY/2017Fall/stat133/stat133-hws-fall17/hw02/data')
library("readr")
dat2 <- read_csv("nba2017-player-statistics.csv", col_types = cols(
  Player= col_character(),
  Team = col_character(),
  Position = col_factor(c("C", "PF", "PG", "SF", "SG")),
  Experience = col_character(),
  Salary = col_double(),
  Rank = col_integer(),
  Age = col_integer(),
  GP = col_integer(),
  GS = col_integer(),
  MIN = col_integer(),
  FGM = col_integer(),
  FGA= col_integer(),
  Points3 = col_integer(),
  Points3_atts= col_integer(),
  Points2 = col_integer(),
  Points2_atts = col_integer(),
  FTM = col_integer(),
  FTA = col_integer(),
  OREB = col_integer(),
  DREB = col_integer(),
  AST = col_integer(),
  STL = col_integer(),
  BLK = col_integer(),
  TO = col_integer()
))



dat2
```


```{r}
str(dat2)
```


#3. Right after Importing the Data
```{r}
dat$Experience[dat$Experience=="R"]<- "0" #convert R to 0
dat$Experience <- as.integer(dat$Experience)
dat
```

#4. Performance of Players 
```{r}
dat$Missed_FG = dat$FGA - dat$FGM
dat$Missed_FT = dat$FTA - dat$FTM
dat$PTS = 3*dat$Points3 + 2*dat$Points2 + dat$FTM
dat$REB = dat$OREB + dat$DREB
dat$MPG = dat$MIN/ dat$GP


dat$EFT = (dat$PTS + dat$REB + dat$AST + dat$STL + dat$BLK - dat$Missed_FG - dat$Missed_FT - dat$TO)/dat$GP
dat
```


```{R}
summary(dat$EFT)
```
```{r}
hist(dat$EFT, col = "gray", xlab = "EFT", main = "Histogram of Efficiency (EFT)")
```


```{r}
#Display the player name, team, salary, and EFT of the top 10 players in decreasing order
new <- dat[order(dat$EFT, decreasing = TRUE),]
new <- new[1:10,]
new[, c("Player", "Team", "Salary", "EFT")]

```


```{r}
dat[dat$EFT<0,1] #Player with negavite EFT
```


```{r}
corrs <- c(
cor(dat$EFT, dat$PTS),
cor(dat$EFT, dat$REB),
cor(dat$EFT, dat$AST),
cor(dat$EFT, dat$STL),
cor(dat$EFT, dat$BLK),
cor(dat$EFT, -dat$Missed_FG),
cor(dat$EFT, -dat$Missed_FT),
cor(dat$EFT, -dat$TO))
corrs <- sort(corrs, decreasing= TRUE)
corrs
```

```{r}
barplot(corrs, names.arg = c("PTS", "REB", "STL", "AST","BLK", "Missed_FT", "Missed_FG", "TO"), yaxp = c(-1, 1, 4), col = c(rep("gray",5), rep("red",3)), main = "Correlations between Player Stats and EFF")

```


#5 Efficiency and Salary
```{r}
plot(dat$EFT, dat$Salary,  col = "blue", xlab = "Efficiency", ylab = "Salary", main = "Scaterplot of Efficiency and Salary and Lowess lines")
lines(lowess(dat$EFT, dat$Salary))
```


```{r}
cor(dat$EFT, dat$Salary)
```

The two variables, EFT and Salary, are positively correlated with correlation coefficient of 0.655624.



```{r}
players2 <- dat[dat$MPG >= 20,]
plot(players2$EFT, players2$Salary, xlab = "EFT", ylab = "Salary", main = "Players vs. Salary")
lines(lowess(players2$EFT, players2$Salary))
```

```{r}
cor(players2$EFT, players2$Salary)
```

The correlation is lower for more established players. This means that EFT and Salary are not really related to each other. This might be due to the fact that the rising players are seeking to be paid more through new support from outside sources. 

Compared to the first plot, the lowess line of the second plot seems to be a little more steeper since it is less curvy. 


#6. Comments and Reflections

The hardest part was to figure out read_csv using column types, and the easy part was indexing columns of the data frame. It took me 5 hours to complete this. 
It was surprising and unexpected to see that the correlation was lower for more established players. 
