---
title: "hw02"
author: "Sierra Park"
date: "9/25/2017"
output: github_document
---


#1 Create a Data Dictionary 
refer to the dictionary file

#2 Import the Data in R

```{r}
setwd('/Users/spark/Documents/UCBERKELEY/2017Fall/stat133/stat133-hws-fall17/hw02/data')
classes <- c("character", "character", "factor", "character", "double", rep("integer",19))
dat <- read.csv(file = "nba2017-player-statistics.csv", colClasses = classes)
dat
```

```{r}
str(dat)
```


```{r}
setwd('/Users/spark/Documents/UCBERKELEY/2017Fall/stat133/stat133-hws-fall17/hw02/data')
#install.packages("readr")
#classes <- list(col_charcter(), col_character(), col_factor(), col_character(), col_double(), rep(col_integer(),19))
library("readr")
dat2 <- read_csv("nba2017-player-statistics.csv", col_types = cols(
  Player= col_character(),
  Team = col_character(),
  Position = col_factor(c("C", "PF", "PG", "SF", "SG")),
  Experience = col_character(),
  Salary = col_double(),
  Rank = col_integer(),
  Age = col_integer(),
  GP = col_integer(),
  GS = col_integer(),
  MIN = col_integer(),
  FGM = col_integer(),
  FGA= col_integer(),
  Points3 = col_integer(),
  Points3_atts= col_integer(),
  Points2 = col_integer(),
  Points2_atts = col_integer(),
  FTM = col_integer(),
  FTA = col_integer(),
  OREB = col_integer(),
  DREB = col_integer(),
  AST = col_integer(),
  STL = col_integer(),
  BLK = col_integer(),
  TO = col_integer()
))



#list(col_character(), col_character(), col_factor(), col_character(), col_double(), rep("col_integer()",19)))
#coltypes(data) <- classes
dat2
```


```{r}
str(dat2)
```


#3. Right after Importing the Data
```{r}
dat$Experience[dat$Experience=="R"]<- "0" #convert R to 0
dat$Experience <- as.integer(dat$Experience)
dat
```

#4. Performance of Players 
```{r}
#PTS: total points => 3*Points3_atts + 2*Points2_atts + FTM
#REB: total rebounds => ORB + DRB
#AST: assists
#STL: steals
#BLK: blocks
#Missed FG: missed field goals => FGA - FGM
#Missed FT: missed free throws => FTA - FTM
#TO: turnovers
#GP: games played 
dat$Missed_FG = dat$FGA - dat$FGM
dat$Missed_FT = dat$FTA - dat$FTM
dat$PTS = 3*dat$Points3 + 2*dat$Points2 + dat$FTM
dat$REB = dat$OREB + dat$DREB
dat$MPG = dat$MIN/ dat$GP


dat$EFT = (dat$PTS + dat$REB + dat$AST + dat$STL + dat$BLK - dat$Missed_FG - dat$Missed_FT - dat$TO)/dat$GP
dat
```


```{R}
summary(dat$EFT)
```

```{r}
#Display the player name, team, salary, and EFT of the top 10 players in decreasing order
new <- dat[order(dat$EFT, decreasing = TRUE),]
new <- new[1:10,]
new[, c("Player", "Team", "Salary", "EFT")]

```


```{r}
dat[dat$EFT<0,1] #Player with negavite EFT
```


```{r}
corrs <- c(
cor(dat$EFT, dat$PTS),
cor(dat$EFT, dat$REB),
cor(dat$EFT, dat$AST),
cor(dat$EFT, dat$STL),
cor(dat$EFT, dat$BLK),
cor(dat$EFT, -dat$Missed_FG),
cor(dat$EFT, -dat$Missed_FT),
cor(dat$EFT, -dat$TO))
corrs <- sort(corrs, decreasing= TRUE)
```

```{r}
barplot(corrs, names.arg = c("PTS", "REB", "STL", "AST","BLK", "Missed_FT", "Missed_FG", "TO"), yaxp = c(-1, 1, 4), col = c(rep("gray",5), rep("red",3)), main = "Correlations between Player Stats and EFF")

```


#5 Efficiency and Salary
```{r}
plot(dat$EFT, dat$Salary,  col = "blue", xlab = "Efficiency", ylab = "Salary", main = "Scaterplot of Efficiency and Salary and Lowess lines")
lines(lowess(dat$EFT, dat$Salary))
```


```{r}
cor(dat$EFT, dat$Salary)
```

The two variables, EFT and Salary, are positively correlated with correlation coefficient of 0.655624.



```{r}
players2 <- dat[dat$MPG >= 20,]
plot(players2$EFT, players2$Salary, xlab = "EFT", ylab = "Salary", main = "Players vs. Salary")
lines(lowess(players2$EFT, players2$Salary))
```

```{r}
cor(players2$EFT, players2$Salary)
```

The correlation is lower for more established players.


#6. Comments and Reflfections

The hardest part was to figure out read_csv using column types, and the easy part was indexing columns of the data frame. It took me 5 hours to complete this. 

