---
title: "hw03"
author: "Sierra Park"
date: "10/5/2017"
output: github_document
---

```{r}
library(dplyr)
library(ggplot2)
```





# Ranking of Teams 

```{r}
getwd()
teams <- read.csv("../data/nba2017-teams.csv", stringsAsFactors = FALSE)
teams
```
## Basic Rankings 

```{r}
descsalary <-  teams %>% arrange(desc(salary))
descsalary
```

```{r}
ggplot(descsalary, aes(x = reorder(team, salary), y = salary))+ geom_bar(stat = 'identity') + coord_flip() + labs(y = "Salary (in Millions)", x = "Team") + ggtitle("NBA Teams ranked by Total Salary") + theme(plot.title = element_text(hjust = 0.5)) + geom_hline(yintercept = mean(descsalary$salary), col = "red")
```



```{r}
ggplot(teams, aes(x = reorder(team, points), y = points))+ geom_bar(stat = 'identity') + coord_flip() + labs(y = "Total Points", x = "Team") + ggtitle("NBA Teams ranked by Total Points") + theme(plot.title = element_text(hjust = 0.5)) + geom_hline(yintercept = mean(descsalary$points), col = "red")
```

```{r}

#EFFICIENCY GRAPH IS WRONG BC EFFICIENCY COLUMN IS WRONG 
ggplot(teams, aes(x = reorder(team, efficiency), y = efficiency))+ geom_bar(stat = 'identity') + coord_flip() + labs(y = "Total Points", x = "Team") + ggtitle("NBA Teams ranked by Total Efficiency") + theme(plot.title = element_text(hjust = 0.5)) + geom_hline(yintercept = mean(descsalary$efficiency), col = "red")
```


According to the rank by total salary, CLE was number one.
According to the rank by total points, GSW was number one.
According to the rank by total efficiency, CLE was number one. 



# Principal Components Analysis (PCA)

```{r}
str(teams[, c(-1, -2, -3, -7, -15)])
```

```{r}
teams[, c(-1, -2, -3, -4, -8, -16)]
```

```{r}
pca <- prcomp(teams[, c(-1, -2, -3, -4, -8, -16)], scale. = TRUE)
```




```{r}
names(pca)
```


```{r}
pca$x
```



```{r}
data.frame(eigenvalue = round(pca$sdev^2, 4), proportion = round(pca$sdev^2 / sum(pca$sdev^2), 4), cumprop = cumsum(round(pca$sdev^2 / sum(pca$sdev^2), 4)))
```


```{r}
scores =as.data.frame(pca$x)
scores$PC1

scores
teams$PC1 <-  scores$PC1
teams$PC2 = scores$PC2
teams
```


```{r}
ggplot(data = teams, aes(x = teams$PC1, y = teams$PC2)) + geom_text(aes(label = teams$team)) + ggtitle(label = "PCA plot (PC1 and PC2)") + labs(x ="PC1", y = "PC2")
```






## Index based on PC1
```{r}
s1 <- (teams$PC1 - min(teams$PC1)) / (max(teams$PC1) - min(teams$PC1)) * 100
teams$scaledPC <- s1
```



```{r}
ggplot(teams, aes(x = reorder(team, scaledPC), y = scaledPC))+ geom_bar(stat = 'identity') + coord_flip() + labs(y = "First PC (scaled from 0 to 100)", x = "Team") + ggtitle("NBA Teams ranked by scaled PC1") + theme(plot.title = element_text(hjust = 0.5))
```

DESCRIPTION:
Ranking by the scaled PC1, GSW seems to rank first by a significant difference. 


# Comments 

Doing PCA was difficult 





